Fractional anisotropy as an optimal tracer of cosmic voids

### 2014-08-12 cosmic_voids.tex
Adding organization of the paper

Here we introduce a new tracer of the structure of cosmic voids built from 
two tensorial web schemes, i.e. the T-web, based on the Hessian of the 
gravitational potential or tidal tensor \citep{Hahn07, Forero09}, and the 
V-web, based on the velocity shear tensor \citep{Hoffman12}. These web 
schemes classify the Cosmic Web into four different types of environment
depending on the counting of the number of eigenvalues below an 
user-defined threshold ($\lambda_{th}$), namely voids (3 eigenvalues below
$\lambda_{th}$), sheets (2), filaments (1) and knots or clusters (0).
The tidal and the shear tensors have proven to be more fundamental than 
the density field as they also trace the collapsing or expanding nature of 
the matter field, which ultimately is what defines the underlying dynamics 
of the Cosmic Web. Furthermore, the density field is degenerated regarding 
the defined types of environment \citep{Hahn07}, which implies that is not 
possible to assign a priori a range of density values to each of them. 
This yields that the usually adopted definition of voids as underdense 
regions in the large-scale matter distribution, is not precise enough as 
it excludes a proper description of the internal structure of voids.

Following the recent trend of using digital image-processing techniques 
developed in other disciplines (e.g. medical sciences and computer imaging) 
for analysing the structure of the Cosmic Web, we propose here, much in 
the same way as \citet{Libeskind13}, the fractional anisotropy (FA) as a 
tracer of the internal structure and the outline of cosmic voids. The FA 
was initially introduced by \citet{Basser95} for quantifying the 
anisotropy degree of the diffusivity of water molecules through cerebral
tissue in nuclear magnetic resonance imaging. In this context we use the 
same quantity for determining the anisotropy degree of the local 
environment from either the orbital dynamics as set by the eigenvalues of
the tidal tensor (T-web), or the dynamics of the velocity field as set by 
the eigenvalues of the shear tensor (V-web). In both cases, the FA is not
determined directly from the density field, hence it is suitable for 
describing the structure of high non-linear regions (e.g. filaments, 
knots and very dense walls) as well as for revealing the fainter 
substructure exhibited by quasi-linear regions like voids.

Once established the FA as the tracer field of cosmic voids, we proceed to
identify individual voids as basins of local minima. For this purpose we
implement a technique defined by \citet{Beucher79} and \citet{Beucher93} 
and widely used in the field of image analysis and mathematical morphology, 
i.e. the \textit{watershed transform algorithm}. This technique has been 
already implemented on void finding schemes by \citet{Platen07} and 
\citet{Neyrinck08} with very interesting results, where individual voids 
are identified as catching basins of local minima of the density field. 
The attractiveness of this algorithm lies in it is parameter-free and does 
not require to make any assumption on the shape and morphology of voids. 
Although we just use a \textit{cloud-in-cell} (\texttt{CIC}) algorithm on 
a Cartesian mesh for estimating the density and tensor fields, instead of 
the more sophisticated \textit{Delaunay tessellation for field estimator} 
(\texttt{DTFE}) technique \citep{Schaap00}, our implementation of the 
watershed transform should not be significantly affected as we are 
interested in quasi-linear regions where the \texttt{CIC} gives similar 
estimations.

This paper is organized as follows.

>>>>

Here we introduce a new tracer of the structure of cosmic voids built from 
two tensorial web schemes, i.e. the T-web, based on the Hessian of the 
gravitational potential or tidal tensor \citep{Hahn07, Forero09}, and the 
V-web, based on the velocity shear tensor \citep{Hoffman12}. These web 
schemes classify the Cosmic Web into four different types of environment
depending on the counting of the number of eigenvalues below an 
user-defined threshold ($\lambda_{th}$), namely voids (3 eigenvalues below
$\lambda_{th}$), sheets (2), filaments (1) and knots or clusters (0).
The tidal and the shear tensors have proven to be more fundamental than 
the density field as they also trace the collapsing or expanding nature of 
the matter field, which ultimately is what defines the underlying dynamics 
of the Cosmic Web. Furthermore, the density field is degenerated regarding 
the defined types of environment \citep{Hahn07}, which implies that is not 
possible to assign a priori a range of density values to each of them. 
This yields that the usually adopted definition of voids as simply 
underdense regions in the large-scale matter distribution, is not precise 
enough as it excludes a proper description of the internal structure of 
voids.

Following the recent trend of using digital image-processing techniques 
developed in other disciplines (e.g. medical sciences and computer imaging) 
for analysing the structure of the Cosmic Web, we propose here, much in 
the same way as \citet{Libeskind13}, the fractional anisotropy (FA) as a 
tracer of the internal structure and the outline of cosmic voids. The FA 
was initially introduced by \citet{Basser95} for quantifying the 
anisotropy degree of the diffusivity of water molecules through cerebral
tissue in nuclear magnetic resonance imaging. In this context we use the 
same quantity for determining the anisotropy degree of the local 
environment from either the orbital dynamics as set by the eigenvalues of
the tidal tensor (T-web), or the dynamics of the velocity field as set by 
the eigenvalues of the shear tensor (V-web). In both cases, the FA is not
determined directly from the density field, hence it is suitable for 
describing both, the structure of high non-linear regions (e.g. filaments,
knots and very dense walls) as well as for revealing the fainter 
substructure exhibited by quasi-linear regions like voids.

Once established the FA as the tracer field of cosmic voids, we proceed to
identify individual voids as basins of local minima. For this purpose we
implement a technique defined by \citet{Beucher79} and \citet{Beucher93} 
and widely used in the field of image analysis and mathematical morphology, 
i.e. the \textit{watershed transform algorithm}. This technique has been 
already implemented on void finding schemes by \citet{Platen07} and 
\citet{Neyrinck08} with very interesting results, where individual voids 
are identified as catching basins of local minima of the density field. 
The appeal of this algorithm lies in that is parameter-free and does not 
require to make any assumption on the shape and morphology of voids. 
Although we use a \textit{cloud-in-cell} (\texttt{CIC}) algorithm on a 
Cartesian mesh for estimating the density and tensor fields, instead of 
the more sophisticated \textit{Delaunay tessellation for field estimator} 
(\texttt{DTFE}) technique \citep{Schaap00}, our implementation of the 
watershed transform should not be significantly affected as we are 
interested in quasi-linear regions where the \texttt{CIC} gives similar 
estimations.

This paper is organized as follows. In section 2 we describe the used
simulation for testing our void finding algorithm, i.e. the Bolshoi
simulation. Alternatively to the web schemes T-web and V-web, we propose 
a very basic web classifier based on three threshold values defined on the
density field, this is used in our comparisons as a representative of all 
density-based schemes. A detailed description of these schemes is 
presented in section 3. In section 4 we describe how the FA can be used 
for tracing the structure of voids. Once obtained meaningful catalogues of 
voids for each scheme, we proceed to compare some typical and general 
interest properties of voids such as volume functions, distributions of 
size and density and velocity profiles. This is done in section 5. Finally
in section 6 we analyse and evaluate our findings.
